%\VignetteEngine{knitr::knitr}
---
title: 'growmod package example'
author: "Mollie Brooks"
date: "16 March 2016"
output: pdf_document
---
```{r}
library(growmod)
library(ggplot2)
```

Simulate data and graph it to get an idea of the general pattern.
```{r, tidy=TRUE, warning=FALSE, message=FALSE}
t1=c("(Intercept)"=1, age=.15,  "I(age^2)"=-.01, prevsize=.7, "prevsize:age"=0, sigma_proc=.01, time_growth_sd=.1, indiv_growth_sd=.1, indiv_age_growth_sd=0, indiv_cor=0, size0_mu=2.5, size0_sd=.5)

set.seed(1111)
sg=simobs(t1, nind=500, ntime=20, sigma_obs=.01, recap=.4)
ggplot(sg, aes(age, size)) + geom_point(alpha=.05) + geom_smooth() + theme_bw()
```

Fit some models of varying complexity.
```{r}
m1=growmod(~1, ~1, data=sg)
m2=growmod(~age, ~1, data=sg)
m3=growmod(~age+I(age^2), ~1, data=sg)
m4=growmod(~age+I(age^2), ~age, data=sg)
m5=growmod(~age, ~age, data=sg)
```

#Model Selection

Now do model selection using either AIC across all models or LRT for pairs of nested models. The AIC value is corrected for small sample sizes.
```{r}
m1$AIC
m2$AIC
m3$AIC
m4$AIC
m5$AIC
LRtest(m2,m1)
LRtest(m3,m2)
LRtest(m4,m3)
LRtest(m4,m5)
```
FIXME: In the future I need to make the models compatible with `AIC` and `AICtab` functions.

Model selection results tell us that `m3` is the best. That's good because it's the form of the true model we used to simulate the data. 

Let's compare the estimates to the true parameter values.
```{r}
summary(m3)
t1
```
The values are pretty close. So that's good.

#Residuals

We can use the `predict` function to make predictions of the next size. These only include the fixed effects. 
```{r, message=FALSE}
dat2=ddply(organize_data(sg, sigma_obs = 0)$Ldat$Predictors, ~ID, mutate,
	prevsize= size[match(age, age+1)]
)
dat2$pred=predict(m3, dat2)
dat2$residuals=dat2$pred-dat2$size
ggplot(dat2, aes(age, residuals))+geom_point()+geom_smooth() + theme_bw()
```